Reproducibility of SNP calling in multiple sequencing runs from single tumors

### 2014-10-28 PAPER/paper_d1.tex
flowcharts

%C) Address potential shortcomings. What could have gone wrong, which parts of the research might be misleading, are there any other caveats?
Although the data is suggestive, it is far from conclusive. In this analysis, we use repeatability between technical overlaps (being in the WGS and WGA samples) as a measure of confidence in a putative SNP. This metric is potentially problematic for two reasons:
%-- using repatability is also problematic 1) cancer is heterogenous; just becuase a mutaiton doesn't show up in the replicate doesn't mean it's not real
(i) First, cancer is highly heterogeneous \cite{hetero}, and so a legitimate somatic SNP might show up in one replicate and not another.
%-- using repatability is also problematic 2) could be degraded DNA, so large number of mutations could still be false positive
(ii) Second, if the DNA sample is degraded to some extent, due to surgery conditions or some other factor out of the hands of the sequencing center, the same errors may appear in both replicates. Although repeatability does not and could not perfectly measure confidence, it should correlate. Having an independent measure of the confidence in SNP calls (repeatability), even an imperfect one, can help us gague the accuracy of other measures (filtering).  
%-- we may have shown that filtering doesn't do very much good, but we haven't really given alternatives
%-- LOH and VAQ each remove a binch of probably-real signal, but they remove a bunch of false signal too
In this sample, we have shown that while some filters function 
%-- other snp callers

%E) (optional) End with a paragraph that again summarizes your research and highlights the most important findings

@@

>>>>

%C) Address potential shortcomings. What could have gone wrong, which parts of the research might be misleading, are there any other caveats?
Although the data is suggestive, it is far from conclusive. In this analysis, we use repeatability between technical overlaps (being in the WGS and WGA samples) as a measure of confidence in a putative SNP. This metric is potentially problematic for two reasons:
%-- using repatability is also problematic 1) cancer is heterogenous; just becuase a mutaiton doesn't show up in the replicate doesn't mean it's not real
(i) First, cancer is highly heterogeneous \cite{hetero}, and so a legitimate somatic SNP might show up in one replicate and not another.
%-- using repatability is also problematic 2) could be degraded DNA, so large number of mutations could still be false positive
(ii) Second, if the DNA sample is degraded to some extent, due to surgery conditions or some other factor out of the hands of the sequencing center, the same errors may appear in both replicates. Although repeatability does not and could not perfectly measure confidence, it should correlate. Having an independent measure of the confidence in SNP calls (repeatability), even an imperfect one, can help us gague the accuracy of other measures (filtering).  
%-- we may have shown that filtering doesn't do very much good, but we haven't really given alternatives
%-- LOH and VAQ each remove a binch of probably-real signal, but they remove a bunch of false signal too
In this sample, we have shown that while some filters remove primarily putative SNPs in only one of two replicates (more likely to be false positives), two filters, VAQ and LOH, remove most of the overlap (more likely to be true positives). While our results suggest that VAQ and LOH filters are probably removing some true positives, that does not preclude them removing false positives. Therefore, if we remove these filters, we may retain more true positives, but we also retain more false positives.
%-- other snp callers
Finally, we looked at results from one SNP caller. There are many more equally widely used SNP callers available, which might do better or worse than the one we have chosen here.

%E) (optional) End with a paragraph that again summarizes your research and highlights the most important findings
We have shown that

@@

\begin{figure}
\includegraphics[scale=1.0]{data_processing_flowchart_2.png}
\caption{\textbf{Number of putative SNPs in WGS v. WGA, as called by SomaticSniper.} Each point is a patient. The line is y=x, so points falling below the line agree with the hypothesis that whole genome amplification makes more mutations in a sample.}
\end{figure}

\begin{figure}
\includegraphics[scale=1.0]{overlap_explained.png}
\caption{\textbf{Number of putative SNPs in WGS v. WGA, as called by SomaticSniper.} Each point is a patient. The line is y=x, so points falling below the line agree with the hypothesis that whole genome amplification makes more mutations in a sample.}
\end{figure}
